---
  title: "VarroaPop MCMC main notebook"
author: "Jeffrey Minucci"
date: "February 20, 2018"
output: 
  html_document:
  toc: yes
---
  
  ```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE)
```
<base target="_top"/>
  <br>
  
  
  Introduction
=====================
  
  Code to infer VarroaPop parameter distributions given bee colony field data featuring neonicotinoid exposure.

  
  Set up for user's computer
====================
```{r}
#Determine path directory based on the user machine######
#jeff epa dev machine
if(Sys.info()[4]=="DZ2626UJMINUCCI"){
vpdir<-path.expand("d:/Git_files/minucci_vp_mcmc/")
vrp_filename <- "default_jeff.vrp"
}

#tom epa windows 2
if(Sys.info()[4]=="DZ2626UTPURUCKE"){
vpdir<-path.expand("k:/git/minucci_vp_mcmc/")
vrp_filename <- "default_tom.vrp" #will need to be generated from default_jeff.vrp with pointer to Tom's weather file location
}

```


Set some VP initial parameters
============================
```{r}
#simulation start and end
#must have mm/dd/yyyy format
SimEnd <- "8/25/2015"
IPollenLoad <- 15
INectarLoad <- 30
IPollenTrips <- 7
INectarTrips <- 17
NeedResourcesToLive <- "false"

#static parameter list
static_names <- c("SimEnd","IPollenLoad","INectarLoad","IPollenTrips","INectarTrips", "NeedResourcesToLive")
static_values <- c(SimEnd,IPollenLoad,INectarLoad, IPollenTrips, INectarTrips, NeedResourcesToLive)
static_list <- list(names = static_names, values = static_values)


```


Choose VP parameters to optimize
===================================
```{r}
#parameters to optimize via MCMC
optimize_names <- c("ICQueenStrength","ICForagerLifespan", "AIAdultLD50", "AIAdultSlope",
                    "AILarvaLD50", "AILarvaSlope", "AIHalfLife", "sd")
#   Notes: ICForagerLifespan appears to be converted to integer by removing decimal places in VP
bound_l <- c(1, 4,  .0001, 2, .0001, 2, 0.01, 0) #lower bondary of the domain for each parameter to be optimized
bound_u <- c(5, 16, .05,   9, .05,   9, 50, 100000) #upper bondary of the domain for each parameter to be optimized
#scales <- (bound_u-bound_l)/10 #for now using the range divided by 10
optimize_list <- list(names = optimize_names, bound_l = bound_l, bound_u = bound_u)

```

Run MCMC algorithm
===================================
  
  Run MCMC for VarroaPop, fitting all 10 sites in parallel

```{r}
## Test vp MCMC function
#library(compiler)
source(paste(vpdir,"src/00run_vp_mcmc.R",sep = "")) 
system.time(test_results <- new_vp_mcmc_c(nsims = 20000, step_length = .15, vp_dir=vpdir, static_vars = static_list, 
                                          optimize_vars = optimize_list, logs = F, verbose = F, debug=F))[3]
#str(test_results)
print(paste("Acceptance rate:", test_results$accept_rate))
#saveRDS(test_results,file=paste(vpdir,"output/mcmc_run_objects/run_5_3_18.rds",sep = ""))

```

Visualize results of test run
====================================
  Acceptance rate (should be ~.25)
```{r}
test_results <- readRDS(file=paste(vpdir,"output/mcmc_run_objects/run_5_11_18_p4.rds",sep = ""))
print(paste("Acceptance rate:", test_results$accept_rate))
```

Plot likelihood trace
```{r}
plot(test_results$like_trace, type='l')
```


Plot parameter traces and posteriors 
```{r}
library(MCMCvis)
MCMCtrace(as.matrix(test_results$param_trace[, test_results$mcmc_params$optimize_vars[["names"]]]),filename="test3_120k_all",
          iter=120000,wd=paste(vpdir,"reports/figures/",sep=""))
```


Continuing where we left off in the MCMC run
========================================
  We can also adjust the step length. 
```{r, echo=FALSE,results='hide'}
source(paste(vpdir,"src/00run_vp_mcmc.R",sep = "")) 


#test_results <- new_vp_mcmc_c(nsims = 15000, step_length = .15, vp_dir=vpdir, static_vars = static_list, 
#                                          optimize_vars = optimize_list, logs = F, verbose = F, debug=F)
#saveRDS(test_results,file=paste(vpdir,"output/mcmc_run_objects/run_5_11_18_p2.rds",sep = ""))

for(i in 1:20){
  test_results <- continue_vp_mcmc(5000,test_results,step_length = .25)
  saveRDS(test_results,file=paste(vpdir,"output/mcmc_run_objects/run_5_11_18_p4.rds",sep = ""))
}

test_results <- continue_vp_mcmc(10,test_results,step_length = .2)
#saveRDS(test_results,file=paste(vpdir,"output/mcmc_run_objects/run_3_28_18_p9.rds",sep = ""))
test_results <- continue_vp_mcmc(30000,test_results,step_length = .2)
#saveRDS(test_results,file=paste(vpdir,"output/mcmc_run_objects/run_3_28_18_p9.rds",sep = ""))

test_results <- continue_vp_mcmc(100000,test_results,step_length = .2)
#saveRDS(test_results,file=paste(vpdir,"output/mcmc_run_objects/run_3_28_18_p9.rds",sep = ""))


test_results <- continue_vp_mcmc(100000,test_results,step_length = .2)
#saveRDS(test_results,file=paste(vpdir,"output/mcmc_run_objects/run_3_28_18_p9.rds",sep = ""))

```

Visualize results
```{r}
test_results$accept_rate
test_results$bounds_rate

true_accept<- (nrow(test_results$param_trace)*test_results$accept_rate) /
  ((1-test_results$bounds_rate)*nrow(test_results$param_trace))
true_accept

plot(test_results$like_trace, type='l')
plot(test_results$param_trace$ICForagerLifespan, type='l')
plot(test_results$param_trace$ICQueenStrength, type='l')
par(mfrow=c(1,2))
hist(test_results$param_trace$ICForagerLifespan)
hist(test_results$param_trace$ICQueenStrength)
par(mfrow=c(1,1))
MCMCtrace(as.matrix(test_results$param_trace[, test_results$mcmc_params$optimize_vars[["names"]]]),filename="test_575k_full",
          iter=575000,wd=paste(vpdir,"reports/figures/",sep="")) #creates pdf in reports/figures folder


```


Diagnostics using the coda package (and others)
```{r}
library(coda)
library(fitR)
library(MCMCvis)
results <- mcmc(test_results$param_trace[, test_results$mcmc_params$optimize_vars[["names"]]])
effectiveSize(results) 
geweke.diag(results)
geweke.plot(results,ask=T)
heidel.diag(results)
last_40 <- mcmc(results[45000:95000,],start=45000)
heidel.diag(last_40)
autocorr.plot(results) 

#estimate burn in period
plotESSBurn(results)
results_burned <- burnAndThin(results, burn = 20000)
plot(results_burned)
heidel.diag(results_burned)
autocorr.plot(results_burned) 

```

```{r}
summary(results_burned)
results_bt <- burnAndThin(results,burn=50000,thin=100)
results_bt2 <- burnAndThin(results,burn=50000,thin=1000)
results_bt3 <- burnAndThin(results,burn=50000,thin=1000)
#results_bt2 <- burnAndThin(results,burn=35000, thin = 1000)
summary(results_bt)
autocorr.plot(results_bt)
autocorr.plot(results_bt2)
autocorr.plot(results_bt3)
heidel.diag(results_bt)
#heidel.diag(results_bt2)
plot(results_bt)
raftery.diag(results, q=0.5, r=.1,s=0.95,converge.eps=0.001)
raftery.diag(results_burned, q=0.5, r=.1,s=0.95,converge.eps=0.001)
raftery.diag(results_bt, q=0.5, r=0.1,s=0.95,converge.eps=0.001)
#http://www.cryoung.org/wp-content/uploads/2014/09/MCMC-workshop-MBARI-2006.pdf

geweke.diag(results_bt2)
geweke.diag(results_bt3)


MCMCtrace(results,filename="test4_265k",
          iter=65000,wd=paste(vpdir,"reports/figures/",sep="")) #creates pdf in reports/figures folder

MCMCtrace(results_bt,filename="test4_265k_thin100",
          iter=65000,wd=paste(vpdir,"reports/figures/",sep="")) #creates pdf in reports/figures folder


MCMCtrace(results_bt2,filename="test4_290k_thin1k_burn50k",
          iter=290000,wd=paste(vpdir,"reports/figures/",sep="")) #creates pdf in reports/figures folder


```

```{r}
#get maximum likelihood parameter estimates
test_results$param_trace[which(test_results$like_trace == max(test_results$like_trace)),][1,]

#plot mle fit to data


#get 95% c.i. for AILarvaLD50
quantile(test_results$param_trace$AILarvaLD50[50000:length(test_results$like_trace)], c(.05,.95))

```

```{r}

vp_field_data <- paste(vpdir,"data/raw/field_bee_areas.csv",sep="")
field_data <- read.csv(vp_field_data)
bees_per_cm2 <- 1.45  #convert area of bees to individuals  
bee_pops <- as.matrix(field_data[,c("bees_cm2_5","bees_cm2_6","bees_cm2_8")]) * bees_per_cm2 
bee_initial <- round(field_data[,c("bees_cm2_4")] * bees_per_cm2,0)
bee_pops <- cbind(bee_initial,bee_pops)

#fitted mle
fitted <- test_results$prop_out_trace[,,which(test_results$like_trace == max(test_results$like_trace))[1]]
fitted <- cbind(bee_initial,fitted)
fitted

#sd
sd_mle <- test_results$param_trac[which(test_results$like_trace == max(test_results$like_trace))[1],"sd"][1]

pdf(paste(vpdir,"/reports/figures/mle_pop_fitted.pdf",sep=""),w=8,h=14)
par(mfrow=c(5,2))
for(i in 1:10){
  plot(c(4,5,6,8),bee_pops[i,]/1000,type='p', pch=19,ylim=c(0,45),xlab="Month of year",ylab="# of bees (1000s)")
  lines(c(4,5,6,8),fitted[i,]/1000,col='darkgrey',lwd=2)
  lines(c(4,5,6,8),(fitted[i,]+sd_mle)/1000,col='darkgrey',lwd=2,lty=2)
  lines(c(4,5,6,8),(fitted[i,]-sd_mle)/1000,col='darkgrey',lwd=2,lty=2)
}
dev.off()

```