---
title: "VarroaPop MCMC initial testing"
author: "Jeffrey Minucci"
date: "February 20, 2018"
output: 
 html_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE)
```
<base target="_top"/>
<br>


Introduction
=====================

Code for initial testing of the VarroaPop MCMC algorithm

### Current bugs:
* Error if only one (or no) static parameters are included

### To implement:
* Neonic exposure events for each site
* Exact date matching
* Resolve the hive that split


Set up for user's computer
====================
```{r}
#Determine path directory based on the user machine######
#jeff epa dev machine
if(Sys.info()[4]=="DZ2626UJMINUCCI"){
  vpdir<-path.expand("d:/Git_files/minucci_vp_mcmc/")
  vrp_filename <- "default_jeff.vrp"
}

#tom epa windows 2
if(Sys.info()[4]=="DZ2626UTPURUCKE"){
  vpdir<-path.expand("k:/git/minucci_vp_mcmc/")
  vrp_filename <- "default_tom.vrp" #will need to be generated from default_jeff.vrp with pointer to Tom's weather file location
}

```


Set some VP initial parameters
============================
```{r}
#simulation start and end
#must have mm/dd/yyyy format
SimEnd <- "8/25/2015"
koc <- 60
kow <- 2

#static parameter list
static_names <- c("SimEnd","koc","kow")
static_values <- c(SimEnd,koc,kow)
static_list <- list(names = static_names, values = static_values)


```


Choose VP parameters to optimize
===================================
```{r}
#parameters to optimize via MCMC
optimize_names <- c("ICQueenStrength","IPollenTrips","INectarTrips",
                   "ICForagerLifespan")
#   Notes: ICForagerLifespan appears to be converted to integer by removing decimal places in VP
bound_l <- c(1,4,4,4) #lower bondary of the domain for each parameter to be optimized
bound_u <- c(5,30,48,16) #upper bondary of the domain for each parameter to be optimized
#scales <- (bound_u-bound_l)/10 #for now using the range divided by 10
optimize_list <- list(names = optimize_names, bound_l = bound_l, bound_u = bound_u)

```

Run MCMC algorithm
===================================

Run MCMC for VarroaPop, fitting all 10 sites in parallel

```{r}
## Test vp MCMC function
#library(compiler)
source(paste(vpdir,"src/00run_vp_mcmc.R",sep = "")) 
system.time(test_results <- new_vp_mcmc_c(nsims = 20, step_length = .5, vp_dir=vpdir, static_vars = static_list, 
                            optimize_vars = optimize_list, logs = T, verbose = T))[3]
str(test_results)
print(paste("Acceptance rate:", test_results$accept_rate))
#saveRDS(test_results,file=paste(vpdir_output,"mcmc_run_objects/test_2_20_18.rds",sep = ""))

```

Visualize results of test run
====================================
Acceptance rate (should be ~.25)
```{r}
test_results <- readRDS(file=paste(vpdir_output,"mcmc_run_objects/test_2_20_18.rds",sep = ""))
print(paste("Acceptance rate:", test_results$accept_rate))
```

Plot likelihood trace
```{r}
plot(test_results$like_trace, type='l')
```


Plot parameter traces and posteriors 
```{r}
library(MCMCvis)
MCMCtrace(as.matrix(test_results$param_trace[, test_results$mcmc_params$optimize_vars[["names"]]]),filename="test_fulltrace",
          iter=5000,wd=paste(vpdir,"reports/figures/",sep=""))
```


Continuing where we left off in the MCMC run
========================================
We can also adjust the step length. 
```{r}

test_results <- continue_vp_mcmc(20,test_results,step_length = .6)
plot(test_results$like_trace,type='l')

```